---
title: "Hypothesis testing"
output: html_document
---

### T-test

1. Two samples come from same population
2. A sample is drawn from a probability distribution
* Null hypothesis: $\mu$ = $\mu_{0}$
* statistic: t 
* assumption: Normal or almost normal distribution for sample
* 1. Equal sample size? 2. Equal variance? for Welch t-test 

```{r}
#t-test
a = 1:10
b = 7:20
t.test(a,b)
```

### Shapiro-Wilk Test
* test sample normality
* Assumption: samll sample (N<50)

```{r}
shapiro.test(rnorm(100, mean = 5, sd = 3))
shapiro.test(runif(100, min = 2, max = 4))
```

### Chi square test
* test sample normality
* assumption: large sample size (N>30)
```{r}
library(fBasics)
pchiTest(rnorm(500, mean = 5, sd = 3), title = NULL, description = NULL)
pchiTest(runif(500, min = 2, max = 4), title = NULL, description = NULL)
```


### Mann-Whitney test (Non-parametric test)
* does not assume data is drawn from any particular distribution
* Comparing the medians of two samples
* "a non-parametric version of t-test"
```{r}
x <- c(0.80, 0.83, 1.89, 1.04, 1.45, 1.38, 1.91, 1.64, 0.73, 1.46)
y <- c(1.15, 0.88, 0.90, 0.74, 1.21)
wilcox.test(x, y, alternative = "g")        # greater
wilcox.test(x, y, alternative = "greater", 
            exact = FALSE, correct = FALSE) # H&W large sample #approximation
```

### Kolmogorov-Smirnov
* compare whether two sample come from same distribution
* qq plot can help visulization
```{r}
x <- rnorm(50)
x2 <- rnorm(50, -1)
plot(ecdf(x), xlim = range(c(x, x2)))
plot(ecdf(x2), add = TRUE, lty = "dashed")
t.test(x, x2, alternative = "g")
wilcox.test(x, x2, alternative = "g")
ks.test(x, x2, alternative = "l")
```